---
title: "Use API data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use API data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is a simple guide to access data via API. Regional granularity levels includes:

- Enumeration areas (*grunnkretser*)
- Cities (*bydeler*)
- Municipalities (*kommuner*)
- Counties (*fylker*)

# Download geo codes 

All functions that have `get_` prefix access data directly via [SSB
API](https://data.ssb.no/api/klass/v1/api-guide.html). This is the simple guide
to use *norgeo* API functions to get codes for municipalities with selected time
period:

```{r api1, eval=FALSE}
library(norgeo)
dt <- get_code("kommune", from = 2015, to = 2021)
# only current geo code
dt <- get_code("kommune", from = 2021)
```

# Get geo changes

You could download geo code changes via `get_change()` function. The example below
will download municipalities codes that have changed from 2015 to 2021.

```{r api2, eval=FALSE}
dtc <- get_change("kommune", from = 2015, to = 2021)
```

# Track geo changes

To track all municipalities codes that have changed since 2015 to their **current**
codes as in 2021 can be done as follows:

```{r api3, eval=FALSE}
dtt <- track_change("kommune", from = 2015, to = 2021)
```

The difference between `get_change()` and `track_change()` is that, the former will
only download codes that have changed while the later will, in addition, track
these changes to the current codes.

You can also track geo codes that are *split* or *merged* to a new geo codes with
function `track_split()` or `track_merge()`.
